---
title: "p8105_hw3_mh4588.Rmd"
author: "Maggie Hsu"
date: "`r Sys.Date()`"
output: github_document
---
### Data Exploration
```{r load data and packages}
library(tidyverse) #Import the Tidyverse library and additional libraries
library(p8105.datasets)
library(patchwork)
#library(hexbin)

data("ny_noaa") #Load NOAA data from the datasets library
```
```
The dataset "ny_noaa" consists of `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` variables, which are: `r names(ny_noaa)`. These variables describe each observation through its ID, date observed, and the levels of precipitation, snowfall, snow depth, and maximum/minimum temperatures. From the summary statistics of this dataset, missing data is an issue for the weather variables (precipitation, snowfall, snow depth, and temperature extremes) due to the amount of NAs present.


### Data Cleaning and Observations
```{r noaa cleaning}

#Created separate variables for year, month, and day and also converted minimum/maximum temperature variables into numeric values.
ny_noaa = 
  mutate(
    ny_noaa,
    date = ymd(date),
    year = year(date), 
    month = month(date), 
    day = day(date),
    tmax=as.numeric(tmax), 
    tmin = as.numeric(tmin)
    ) 

#Most commonly observed values for snowfall 
ny_noaa |>
  count(snow)
```

The most commonly observed value for snow is 0 because most days in New York do not have any snow.


### Average Maximum Temperatures in January and July

```{r tmax tmin}
ny_noaa |>
  group_by(id, year, month) |>
  filter(month %in% c(1, 7)) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE, color = id)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id, color=year)) +
  geom_point() + 
  geom_path() + 
  facet_grid(.~month)
  labs(title="Mean Maximum vs. Minimum Temperatures by Year", x="Year", y="Temperature")
```
Although maximum and minumum temperatures tend to fluctuate over time, the mean minimum temperature shows an increase from 1980 to 2010 while the mean maximum temperature is stable. There also seems to be some extreme observations (such as a very low minimum for 1980 and 2005 and a very low maximum slightly before 1990)


### Distribution of Snowfall and Temperatures
```{r snowfall}
#Plot temperatures
temperatures = ny_noaa |>
  ggplot(aes(x = tmin, y = tmax)) + geom_hex()
  
#Distribution of snowfall values by year
snowfall = ny_noaa |>
  filter(snow > 0, snow < 100) |>
  ggplot(aes(x = as.factor(year), y = snow, fill=year)) +
  geom_boxplot()

temperatures + snowfall
```

# Problem 2
```{r nhanes cleaning}
#Read in datasets
covar <- read.csv("./data/nhanes_covar.csv")
accel <- read.csv("./data/nhanes_accel.csv")

#Clean data
#Rename rows and remove first four blank rows
covar <- janitor::row_to_names(covar, 4) |>
  filter(age >= 21) |> #Exclude participants under 21 years old
  mutate( #Encode variables into appropriate classes
    SEQN = as.numeric(SEQN),
    sex = factor(sex, labels=c("male","female")),
    age = as.numeric(age), 
    BMI = as.numeric(BMI),
    education = factor(education, labels = c("Less than High School","High School Equivalent", "More than High School")),
  )

#Merge frames based on SEQN
nhanes <- left_join(covar,accel, by="SEQN") |>
  #Remove rows with missing demographic values
  drop_na(SEQN,sex,age,BMI,education)
    
#Create a long/tidy version of the nhanes dataframe
nhanes_long <- pivot_longer(nhanes, min1:min1440, names_to = "time", values_to = "mims")

nhanes_long = mutate(nhanes_long, time= str_sub(time, 4, -1))
nhanes_long = mutate(nhanes_long, time= as.numeric(time)) |>
drop_na(SEQN,sex,age,BMI,education, time, mims)
    
```

```{r nhanes table}
#Set up table comparing the number of men and women in each level of education
nhanes |>
  janitor::tabyl(sex, education)

#Visualization
nhanes |>
  ggplot(aes(x = sex, y = age, fill = sex)) + 
    facet_grid(~education) + geom_boxplot()
    
```
Most participants were in the "More than High School" category as compared to the other two educational levels, while the amount of male and female participants were at similar levels, there were many more male participants at the High School level than female participants. 

The median ages for male and female participants for Less than High School are similar, while for High School Equivalent the male participants tended to be younger than the female participants. For More than High School, the male and female participants' age distributions tended younger than for the other groups and had similar medians--although it was slightly lower for female participants. 

```{r nhanes total activity}
#Create a total activity variable by summing up individual minute activity values for all columns
nhanes = mutate(nhanes, activity = rowSums(across(min1:min1440)))

#Plot activity compared to age, sex, and education levels.
nhanes |>
    ggplot(aes(x=age, y=activity)) + 
  facet_grid( .~education) +  
  geom_point(aes(color = sex)) +
  geom_smooth(se = FALSE) +
  labs(title="Daily Activity Level by Age and Education Level", y="Activity (MIMS Units)", x="Age")
```
There is a negative association between activity level and age for all of the observed education levels as the lines of best fit were all downward trending. It is interesting that in "High School Equivalent", activity levels increase until 40 but then decrease, while activity levels slightly increase at age 60 for "Less than High School" but then decrease again.


```{r nhanes 24 hour}
#Three-panel plot showing activity level over time
nhanes_long |>
  ggplot(aes(x=time, y=mims) + 
  facet_grid( .~education) +  
  geom_point(aes(color = sex)) +
  geom_smooth(se = FALSE) +
  labs(title="Daily Activity Over 24 Hours by Age and Education Level", y="Activity (MIMS Units)", x="Time")
```
All of the education levels show a pattern where activity tends to start off at a moderate level, dips slightly before 500 minutes, and then increases shortly after 500 minutes. The activity level then stays relatively level (or slightly decreases for "Less than High School" participants) until roughly 1000 minutes in, where the activity level decreases until at a similar level to the starting activity level. 
Male and female participants had similar activity patterns, although for "More than High School" the male participants showed a spike in activity around 500 minutes in while female participants showed a spike somewhat after 1000 minutes in. 


# Problem 3
```{r citibike}
#Read in data
jan_2020 <- read.csv("./data/citibike/Jan 2020 Citi.csv") 
jan_2024 <- read.csv("./data/citibike/Jan 2020 Citi.csv")
july_2020 <- read.csv("./data/citibike/July 2020 Citi.csv")
july_2024 <- read.csv("./data/citibike/July 2024 Citi.csv")

#Add identifier variables
jan_2020 = mutate(jan_2020, month = "January 2020")
jan_2024 = mutate(jan_2024, month = "January 2024")
july_2020 = mutate(july_2020, month = "July 2020")
july_2024 = mutate(july_2024, month = "July 2024")

#Merge all datasets
citibike <- bind_rows(jan_2020, jan_2024, july_2020, july_2024)
```

```{r citibike table}
citibike |>
  group_by(member_casual, month) |> #Group by member types and summarize
  summarize(total_rides = n()) 
```
There tend to be more citibike rides in July  as compared to in January for both members and casual riders, which may be due to warmer weather over the summer. Members have more overall rides as compared to casual users as members may be more inclined to regularly use citibikes, while there are either the same or more rides in 2024 overall as pandemic restrictions and quarantines in 2020 would have made people less likely to go out or rent bikes.

## The 5 most popular starting stations in July 2024
```{r station table}
citibike |> 
  filter(month=="July 2024") |>
  group_by(start_station_name) |> #Group by station name and number of rides per station
  summarize(originating_rides = n())|>
  filter(min_rank(desc(originating_rides)) < 6) 
```

## Visualizations
### Plot 1
```{r citibike plot 1, fig.width=20}
citibike |>
  group_by(month) |>
  ggplot()+
  geom_boxplot(aes(x=weekdays, y=duration, fill=weekdays))+ 
  facet_grid( .~month) +
  labs(title="Citibike Durations by Days of the Week", x ="Day of the Week", y="Duration")
```
In general, there tended to be more rides on Saturday and Sunday as compared to the weekdays, especially on Tuesdays and Wednesdays. July has a wider range of ride durations and a higher maximum for both 2020 and 2024, suggesting that people may tend to ride bikes more over the summer as compared to January. There are also more long-duration outliers for July rides. 


### Plot 2
```{r citibike plot 2}
citibike |>
  filter(month == c("January 2024","July 2024")) |>
  ggplot(aes(x=month, y=duration, color=member_casual) 
  )+
  geom_boxplot() +
  label(title="Citibike Usage Patterns in 2024")
```

From this graphic, casual users tended to have a higher range of bike durations as compared to members, and had a higher minimum, maximum, and median for these values for both January and July 2024. The duration patterns were generally similar for both January and July, although for outliers, there were higher amounts of high-duration outliers for July 2024 as compared to January. 